---
title: "Challenge 2 Answer Key"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = F)
```

**1. Change** `sex` **and** `ulcer` **to categorical variables using `mutate`.** *(5 points)*
```{r}
Melanoma <- Melanoma %>% 
  mutate(sex = factor(sex, labels = c("Female","Male")), ulcer = factor(ulcer))
```
*1 Point*: Correct labels for `sex`  
*1 Point*: `dplyr` used  
*3 Points*: `sex` and `ulcer` changed to factors  

**2. Add** `died` **variable to the dataset.** *(5 points)*
```{r}
Melanoma <- Melanoma %>% mutate(died = ifelse(status == 1, 1, 0)) 
```
*1 Point*: `dplyr` used    
*4 Points*: `died` added correctly     

**3. Create a new table using the pipe command.** *(5 points)*
```{r}
Melanoma %>% 
  group_by(ulcer) %>%
  filter(age > 65) %>% 
  summarise (n = n(), age = mean(age), thickness = mean(thickness), died = mean(died)) %>%      
  arrange(desc(n))
```
*2 Points*: `group_by` variable changed to `ulcer`    
*1 Point*: Correct age range selected    
*2 Points*: Table created and interpreted correctly   

**4. Create a new subset of the data using** `dplyr`**, then fit a logistic regression model to it using the** `glm` **function.** *(5 points)*
```{r}
Melanoma2 <- Melanoma %>% dplyr::select(sex, age, thickness, ulcer, died) 
mod <- glm(died~., family=binomial(link='logit'), data = Melanoma2)
summary(mod)
```
*1 Point*: `select` used to create new dataframe    
*2 Points*: `glm` function used to fit model     
*2 Points*: Model coefficients interpreted correctly     

*If warnings or messages are included in the RMarkdown output, deduct one point from the final score.*
